Locally-Oriented Programming: A simple programming model for stencil computations

### 2015-01-22 papers/LOPe/local.tex
Starting work on EuroPar version.

\title{Locally-Oriented Programming: A stunningly simple programming model for GPU programming in Fortran and C++}

@@

\author{
% 1st. author
\alignauthor
Craig E Rasmussen\\
       \affaddr{Los Alamos National Laboratory}\\
%       \affaddr{CCS-7, MS B287}\\
       \affaddr{Los Alamos, NM 87545}\\
       \email{crasmussen@lanl.gov}
% 2nd. author
\alignauthor
Matthew J. Sottile\\
       \affaddr{Galois, Inc.}\\
%       \affaddr{421 SW 6th Ave. Suite 300}\\
       \affaddr{Portland, OR 97204}\\
       \email{mjsottile@computer.org}
% 3rd. author
\alignauthor
Dan Nagle\\
       \email{dannagle@verizon.net}
% 4th. author
\alignauthor
Daniel Quinlan\\
       \affaddr{Lawrence Livermore National Laboratory}\\
       \email{dquinlan@llnl.gov}
}

@@

\begin{abstract}
Emerging GPU architectures for high performance computing are well suited to a
data-parallel programming model.  However, programming for these architectures
currently requires learning a new language, e.g., OpenCL or CUDA, or employing
OpenMP compiler directives.  This paper presents preliminary work examining how
programming with a local orientation can be employed in Fortran (or using a
special array library in C++) to provide simple access GPU architectures.  A
locally-oriented programming model is especially useful for stencil-based codes
or algorithms requiring the application of convolution kernels.  In this
programming model, a programmer codes the algorithm {\it only} for a single
array element, but has read-only access to a small sub-array surrounding the
given array element, so that a stencil can be applied.  We demonstrate how a
locally-oriented programming model can be adopted using source-to-source program
transformations in conjuction with a compiler-based infrastructure like ROSE.
\end{abstract}

@@

\input{intro}
\input{programming-model}
\input{shallow-water}
\input{source-to-source}
%\input{static-analysis}
\input{performance}

@@

The sheer complexity of programming for clusters of many or multi-core
processors with tens of millions threads of execution makes the simplicity of
the data-parallel model attractive.  The increasing complexity of
today's applications (especially in light of the increasing complexity
of the hardware) and the need for portability across architectures
make a higher-level and simpler programming model like data-parallel
attractive.

The goal of this work has been to exploit source-to-source transformations that
allow programmers to develop and maintain programs at a high-level of
abstraction, without coding to a specific hardware architecture.
Furthermore these transformations allow multiple hardware architectures
to be targeted without changing the high-level source.  It also removes the
necessity for application programmers to understand details of the accelerator
architecture or to know OpenCL.

@@

\bibliographystyle{abbrv}
\bibliography{local}

>>>>

\title{Locally-Oriented Programming: A simple programming model for stencil computations}

@@

\author{
% 1st. author
\alignauthor
Craig E Rasmussen\\
       \affaddr{Los Alamos National Laboratory}\\
%       \affaddr{CCS-7, MS B287}\\
       \affaddr{Los Alamos, NM 87545}\\
       \email{crasmussen@lanl.gov}
% 2nd. author
\alignauthor
Matthew J. Sottile\\
       \affaddr{Galois, Inc.}\\
%       \affaddr{421 SW 6th Ave. Suite 300}\\
       \affaddr{Portland, OR 97204}\\
       \email{mjsottile@computer.org}
% 3rd. author
\alignauthor
Dan Nagle\\
       \email{dannagle@verizon.net}
% 4th. author
\alignauthor
Soren Rasmussen\\
       \affaddr{Utah State University}\\
       \email{soren.rasmussen@aggiemail.usu.edu}
}

@@

\begin{abstract}
Some stuff...

%%Emerging GPU architectures for high performance computing are well suited to a data-parallel programming model.  However, programming for these architectures currently requires learning a new language, e.g., OpenCL or CUDA, or employing OpenMP compiler directives.  This paper presents preliminary work examining how programming with a local orientation can be employed in Fortran (or using a special array library in C++) to provide simple access GPU architectures.  A locally-oriented programming model is especially useful for stencil-based codes or algorithms requiring the application of convolution kernels.  In this programming model, a programmer codes the algorithm {\it only} for a single array element, but has read-only access to a small sub-array surrounding the given array element, so that a stencil can be applied.  We demonstrate how a locally-oriented programming model can be adopted using source-to-source program transformations in conjuction with a compiler-based infrastructure like ROSE.
\end{abstract}

@@

%\input{intro}
%\input{motivation}
\input{programming-model}
%\input{program-restructuring}
%\input{static-analysis}
%\input{performance}

@@

Some stuff...

%%The sheer complexity of programming for clusters of many or multi-core processors with tens of millions threads of execution makes the simplicity of the data-parallel model attractive.  The increasing complexity of today's applications (especially in light of the increasing complexity of the hardware) and the need for portability across architectures make a higher-level and simpler programming model like data-parallel attractive.

%%The goal of this work has been to exploit source-to-source transformations that allow programmers to develop and maintain programs at a high-level of abstraction, without coding to a specific hardware architecture.  Furthermore these transformations allow multiple hardware architectures to be targeted without changing the high-level source.  It also removes the necessity for application programmers to understand details of the accelerator architecture or to know OpenCL.

@@

%\bibliographystyle{abbrv}
%\bibliography{local}
